SunPy - Python for Solar Physics

### 2014-01-29 2-1-Map.tex

\subsection{Map}\label{ssec:map}
The Map data type stores 2D spatial data, such as images of the Sun and 
inner heliosphere. It provides a wrapper around a \texttt{numpy} data array and 
its associated spatial coordinates and other metadata. The \texttt{Map} class 
provides methods for typical operations on 2D data, such as rotation and 
re-sampling, as well as visualisation functionality.
The \texttt{Map} class also provides a convenient interface for loading data 
from a variety of sources, including a FITS file as shown in 
Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} class consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around an \texttt{ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration of \texttt{Astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependant upon the data provided. 
SunPy v0.4 has \texttt{Map} specialisations for the following instruments: 
\textit{YOHKOH/SXT}, \textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames. 
% RJH: Instruments should be consistent throughout (italics, caps, etc).  Also,
% usually citations are provided for instruments.

The \texttt{Map} class stores standard metadata retrieved from the header of 
the image file in the \texttt{meta} attribute and provides convenience 
properties for commonly accessed metadata: \textit{e.g.}, \texttt{instrument}, 
\texttt{wavelength} or \texttt{coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-look functionality of 
\texttt{Map}; see Section \ref{subsec:Viz} for more discussion of plotting.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The map is created from an \textit{AIA} FITS file,
and a quick-view plot is created.}
\label{code:aia_1}
\end{listing}

In addition to the data-type classes, the \texttt{map} sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listings~\ref{code:compmap_1} and \ref{code:mapcube_1} show how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} shows how the \texttt{MapCube} 
classes interact with the \texttt{matplotlib.animation} submodule to save a 
video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example showing a \texttt{CompositeMap} plot, with RHESSI data composed with SDO/AIA data, and the integration with the \texttt{matplotlib.pyplot} interface.}
\label{code:compmap_1}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> from matplotlib import animation
>>> mapc = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> anim = mapc.plot()
>>> Writer = animation.writers['ffmpeg']
>>> writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
>>> anim.save('aia_cube.ogv')
\end{minted}
% In the above example, is Writer, writer clear? 
% Could we add more metadata to the movie? obs_date, instrument, etc...?
\caption{Example showing how to save a video animation from a \texttt{MapCube}, using 
\texttt{matplotlib}'s animation framework.}
\label{code:mapcube_2}
\end{listing}
% RJH: If this is kept, the 'artist' bit should be explained.

>>>>

\subsection{Map}
The Map data type represents 2D spatial data such as images of the Sun and 
inner heliosphere. It provides a wrapper around a data array and its associated
spatial coordinates and other metadata. The \texttt{Map} class provides 
methods for typical operations on 2D data, 
such as rotation and re-sampling, and visualisation functionality.
The \texttt{Map} class
also provides a convenient interface for loading data from a variety of sources, typically
a FITS file as shown in Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} class consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. \texttt{NDData} is a 
generic wrapper around an \texttt{ndarray} with a \texttt{.meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does not yet make full
use of its capabilities, but this inheritance structure provides 
for future integration of \texttt{Astropy}. In order to provide instrument- or
detector-specific integration, \texttt{GenericMap} is subclassed on creation. 
Each subclass of \texttt{GenericMap} can register with the \texttt{Map} creation function (\textit{i.e.}, factory)
by implementing a method that returns \texttt{True} if the metadata
in the file header matches
that instrument or detector.  This \texttt{Map} factory 
will then automatically return an instance of the specific \texttt{GenericMap} 
subclass. SunPy 0.4 has \texttt{Map} specialisations for the 
following instruments: 
\textit{YOHKOH/SXT}, \textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames.

The \texttt{Map} class stores standard metadata retrieved from the header of 
the image file in the \texttt{.meta} attribute and provides convenience 
properties for commonly accessed metadata: \textit{e.g.}, \texttt{.instrument}, 
\texttt{.wavelength} or \texttt{.coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-plot functionality of 
\texttt{Map}; see Section \ref{subsec:Viz} for more discussion of plotting.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The Map is created from an \textit{AIA} FITS file,
and a quick-view plot is created.}
\label{code:aia_1}
\end{listing}

In addition to the data-type classes, the \texttt{map} sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listings~\ref{code:compmap_1} and \ref{code:mapcube_1} show how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} shows how the \texttt{MapCube} 
classes interacts with the \texttt{animation} module to save a video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example showing a \texttt{CompositeMap} plot and the integration with the \texttt{matplotlib.pyplot} interface.}
\label{code:compmap_1}
\end{listing}

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> from matplotlib import animation
>>> mapc = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> anim = mapc.plot()
>>> Writer = animation.writers['ffmpeg']
>>> writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
>>> anim.save('aia_cube.ogv')
\end{minted}
\caption{Example showing how to save a video animation from a \texttt{MapCube}, using 
\texttt{matplotlib}'s animation framework.}
\label{code:mapcube_2}
\end{listing}

### 2014-01-29 2-2-Lightcurve.tex
comments and some little fixings

Time series data and their analysis are a fundamental aspect of solar
physics for which many data sources are available.
SunPy provides a \texttt{LightCurve} class
with a convenient and consistent interface for handling solar time-series
data.  The main engine behind the \texttt{LightCurve} class is
the \href{http://pandas.pydata.org}{\texttt{pandas}} data-analysis library, and 
\texttt{LightCurve}'s \texttt{.data} attribute is a \texttt{pandas.DataFrame} 
class.
The \texttt{pandas} library contains a large amount
of functionality for manipulating and analysing time-series data,
making it an ideal basis for \texttt{LightCurve}.  \texttt{LightCurve}
assumes that the input data are time-ordered list(s) of numbers, and each
list becomes a column in the \texttt{pandas} DataFrame class.

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products, as well as implementing a factory interface 
similar to that of \texttt{Map}.  Since there is no established standard
as to how time-series data should be stored and distributed, each SunPy 
\texttt{LightCurve} object sub-class provides the ability to download its corresponding 
specific data format in its constructor and parse that file type.

A \texttt{LightCurve} object may be created using a number of different methods. 
Firstly, a \texttt{LightCurve} may be created for a specific instrument given
an input time range. In Listing~\ref{code:goes_lc}, 
the \texttt{LightCurve} constructor searches a remote source for the GOES X-ray 
data specified by the time interval, downloads the required files and 
subsequently creates and plots the object.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 08:00')
>>> goes.peek()
%removed the resample part of the example - really we need to fix truncate!
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 
06:00--08:00 UT on 2011 June 7 using a time range, and the output of the 
\texttt{peek()} method.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

@@

%Stuart: Where did the pandas section go?
%ARI - looks like it's up top

>>>>

Time series data and their analysis are a fundamental aspect of solar
physics for which many data sources are available.
SunPy provides a \texttt{LightCurve} class
with a convenient and consistent interface for handling solar time-series
data.  The main engine behind the \texttt{LightCurve} class is
the \texttt{pandas} data-analysis library (\url{http://pandas.pydata.org}),
and \texttt{LightCurve}'s
\texttt{.data} attribute is a \texttt{pandas.DataFrame} class.
The \texttt{pandas} library contains a large amount
of functionality for manipulating and analysing time-series data,
making it an ideal basis for \texttt{LightCurve}.  \texttt{LightCurve}
assumes that the input data are time-ordered list(s) of numbers, and each
list becomes a column in the \texttt{pandas} DataFrame class.

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products.  Since there is no established standard
as to how time-series data
should be stored and distributed, each SunPy \texttt{LightCurve}
object provides the ability to download its corresponding specific
data format in its constructor.

%\subsubsection{Creating and using Lightcurves}

A \texttt{LightCurve} object may be created using a number of different methods. 
Firstly, a \texttt{LightCurve} may be created for a specific instrument given
an input time range. In Listing~\ref{code:goes_lc}, 
the \texttt{LightCurve} constructor searches a remote site for the GOES X-ray data 
specified by the time interval, downloads the required files and subsequently 
creates the object.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 10:00')
>>> goes.peek()
>>> trange = TimeRange('2011-06-07 06:00','2011-06-07 07:00')
%>>> goes_zoom = goes.truncate(trange)
% the above does not work!
>>> resampled_goes = goes.data.resample("1H")
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 06:00--08:00 UT on 
2011 June 7 using a time range, and the output of the \texttt{peek()} method.
A sub-interval is then extracted from the Lightcurve and finally the data is resampled
to 1 hour cadence.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

@@

### 2014-01-29 2-3-Spectra.tex

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...

SunPy aims to provide broad support for solar spectroscopy
instruments.  The variety and complexity of these instruments and
their resultant datasets makes this a challenging goal.  The \texttt{spectra} module implements a
\texttt{Spectrum} class for 1D data (intensity versus frequency) and a
\texttt{Spectrogram} class for 2D data (intensity versus time and
frequency).  Each of these classes use a \texttt{numpy.ndarray} class
as its \texttt{.data} attribute.  These two classes were implemented
by funding provided by the Astrophysics Research Group at Trinity
College Dublin, Ireland.

The \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities.
%The common functionality provided by the base \texttt{Spectrogram} class includes
%reading data,
%joining different time ranges and frequencies,
%performing frequency-dependent background subtraction,
%and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{.from\_range} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
%ji to DPS - what is the linearisation that is happening here?

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.spectra.sources.callisto import CallistoSpectrogram
>>> tstart, tend = "2011-06-07T06:00:00", "2011-06-07T07:45:00"
>>> callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
>>> callisto_nobg = callisto.subtract_bg()
>>> callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the
  data for the requested time range and observatory, merges it and
  removes the background signal.  The data requested -- `BIR' -- is
  the code name of the \href{http://www.rosseobservatory.ie}{Rosse Observatory}
  at Birr Castle in Ireland.}
\label{code:spectra}
\end{listing}

>>>>

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum

%ayshih - this whole paragraph can be deleted
%ji - I agree, this is redundant in such a journal.
Spectroscopy is the study of radiative energy as a function of wavelength.
An spectrum is normally obtained by observing a range of frequencies at the same time; 
it can be measured for example in radio wavelengths by a frequency sweep with 
a radio receiver or in visible and ultraviolet by the dispersion of the incident light 
through a diffraction grating or prism.
The analysis of the resultant spectrum can provide properties of the plasma observed 
such temperature, density, speed, etc.
Therefore, spectroscopy provides to the solar physicists invaluable information about 
the composition and the physical properties of the Sun.

SunPy aims to provide broad support for solar spectroscopy
instruments.  The variety and complexity of these instruments and
their resultant datasets makes it very challenging to support all in a
general way.  The \texttt{spectra} module implements a
\texttt{Spectrum} class for 1D data (intensity versus frequency) and a
\texttt{Spectrogram} class for 2D data (intensity versus time and
frequency).  Each of these classes use a \texttt{numpy.ndarray} class
as its \texttt{.data} attribute.  These two classes were implemented
by funding provided by the Astrophysics Research Group at Trinity
College Dublin, Ireland.

The \texttt{Spectrogram} class supports radio spectrograms from the e-Callisto 
solar radio spectrometer network (\url{http://www.e-callisto.org/})
and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities.
%The common functionality provided by the base \texttt{Spectrogram} class includes
%reading data,
%joining different time ranges and frequencies,
%performing frequency-dependent background subtraction,
%and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{.from\_range} function, the object merges all the downloaded
files downloaded into a single spectrogram, across time and frequency.
In the example shown, data is observed in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
%ji to DPS - what is the linearisation that is happening here?

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.spectra.sources.callisto import CallistoSpectrogram
>>> tstart, tend = "2011-06-07T06:00:00", "2011-06-07T07:45:00"
>>> callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
>>> callisto_nobg = callisto.subtract_bg()
>>> callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the
  data for the requested time range and observatory, merges it and
  removes the background signal.  The data requested -- 'BIR' -- is
  the code name of the Rosse Observatory at Birr Castle in Ireland
  (\url{http://www.rosseobservatory.ie})}
\label{code:spectra}
\end{listing}

### 2014-01-29 3-2-HEK.tex

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g. flares and radio bursts) on many different time-scales, length-scales and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK).  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature/event data elsewhere (for example, the GOES flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012} to populate the HEK.  The advantage of the HEK is it 
provides an homogeneous and well-described interface to a large amount of 
feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was also
developed with support from ESA's SOCIS 2011.  The \texttt{hek} module makes 
use of the 
\href{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}{HEK
 API}.
The current list of events maintained by the HEK and their properties can be 
found at \url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import hek
>>> client = hek.HEKClient()
>>> tstart, tend = '2011/08/09 00:00:00', '2011/08/10 00:00:00'
>>> event_type = 'FL' # 'FL' indicates flare
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type))
>>> len(result)
52
\end{minted}
\caption{Example usage of the \texttt{hek} module showing a simple HEK search for solar flares
which occurred on August 9th, 2011.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators. This allows filtering of the HEK entries by the properties of the 
event. As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods} (FRMs). 
Example in Listing~\ref{code:hek:frm} filters the results of the previous 
result to return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type),
...                       hek.attrs.FRM.Name == 'SSW Latest Events')
>>> len(result)
9
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
It is also possible to use the results of a HEK query to download the 
corresponding data from the VSO by using SunPy's \texttt{H2VClient}.
%schriste - this is a weird way to end this section, either show how it is done or 
% do not mention it, i think.
%dps - I agree.
%Stuart: Will do on my next pass

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena (e.g. flares, 
radio bursts) on many 
different time-scales, length-scales and wavelengths. Observations and metadata 
concerning these phenomena are collected at the Heliophysics Event 
Knowledgebase (HEK).  Entries are generated both by automated algorithms and 
human observers.  Some of the information in the HEK reproduces feature/event 
data elsewhere (for example, the GOES flare catalog), and some is generated by 
the Solar Dynamics Observatory Feature Finding Team \cite{martens2012} to
populate the HEK.  The advantage of the HEK is it provides a homogeneous and 
well-described interface to a large 
amount of feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was 
developed with support from the European Space Agency Summer of Code in 
Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
The current list of 
events maintained by the HEK and their properties can be found at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import hek
>>> client = hek.HEKClient()
>>> tstart, tend = '2011/08/09 00:00:00', '2011/08/10 00:00:00'
>>> event_type = 'FL' # 'FL' indicates flare
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type))
>>> len(result)
52
\end{minted}
\caption{Example usage of the \texttt{hek} module showing a simple HEK search for solar flares
which occurred on the 2011/08/09.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods} ( 
FRMs). Example \ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
                      hek.attrs.FRM.Name == 'SSW Latest Events')
>>> len(result)
9
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.
%schriste - this is a weird way to end this section, either show how it is done or 
% do not mention it, i think.

### 2014-01-29 3-5-Database.tex

\subsection{The File Database}\label{ssec:db}

Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organised and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
SunPy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of the Google 
Summer of Code (GSoC) program 2013.

The database sub-package makes use of a database supported by
\href{http://www.sqlalchemy.org}{SQLAlchemy}. This library was chosen
since it supports many SQL dialects. 
If SQLite is selected, the database is stored as a single file which is
created automatically. A server-based database, on the other hand, could be used
by a collaborators who work together on the
same data from different computers: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that can be read via SunPy's 
\texttt{io} subpackage, and direct integration with the \textsc{VSO} 
subpackage is supported.
It is also possible to manually add file or directory entries. The package also provides
a unified data search via the \texttt{fetch()} method, which includes both local files
and files on the \textsc{VSO}. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database, the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favouring, as well as 
querying the full metadata (i.e., FITS header) of each entry.

@@

The \texttt{Database} class connects to a database and allows the user to 
perform operations on it. Listing~\ref{code:database} shows how to connect
to an in-memory database and download data from the \textsc{VSO}. These entries are
automatically added to the database. The function \texttt{len()} is used to get the number of
records. The function \texttt{display\_entries()} displays an iterable of 
database entries in a nicely-formatted \textsc{ASCII} table. The headlines 
correspond to the attributes of the respective database entries.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time('2012-08-05', '2012-08-05 00:00:05'),
...     vso.attrs.Instrument('AIA'))
>>> len(database)
2
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax'])
id observation_time_start wavemin wavemax
-- ---------------------- ------- -------
1  2012-08-05 00:00:01    9.4     9.4    
2  2012-08-05 00:00:02    33.5    33.5   
\end{minted}
\caption{Example usage of the database sub-package.}
\label{code:database}
\end{listing}

>>>>

\subsection{The File Database}
Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organized and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
SunPy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of the 2013 
Google Summer of Code.

The database sub-package makes use of a database supported by
SQLAlchemy(\url{http://www.sqlalchemy.org}). This library was chosen
since it supports many SQL dialects. If SQLite is choosen, the database is stored as a single file which is
created automatically. This solution makes it possible to easily backup, copy and distribute
a database file. A server-based database, on the other hand, could be used by a collaboration 
who work together from different machines on the
same data: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that is retrieved via sunpy from the net
package though currently only the \textsc{VSO} package is supported. 
It also possible to add entries manually by file or directory. The package also provides
a \texttt{fetch} method which provides a unified data search which includes both local files
and files on the \textsc{VSO} are searched. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favoriting.

@@

Using the database package, one uses the \texttt{Database} class to connect
with a database and perform operations on it. Listing~\ref{code:database} shows how to connect
to an in-memory database and download data from the \textsc{VSO}. These entries are
automatically added to the database. The function \texttt{len} is used to get the number of
records. The function \texttt{display\_entries} displays an iterable of database entries
in a nicely-formatted \textsc{ASCII} table. The headlines correspond to the
attributes of the respective database entries.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time((2011, 9, 20, 1), (2011, 9, 20, 2)),
...     vso.attrs.Instrument('RHESSI'))
>>> len(database)
6
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax', 'download_time'])
id observation_time_start wavemin        wavemax           download_time      
-- ---------------------- -------        -------           -------------      
1  2011-09-20 01:09:20    0.413280643067 7.29318781883e-05 2014-01-18 19:21:35

\end{minted}
\caption{Example usage of the database sub-package.}
\label{code:database}
\end{listing}
